\LARGE \bf Approximately Orchestrated Road Traffic Automata\\ a General Traffic Simulation Framework

### 2012-03-07 papers/itsc_2012/draft2.tex
Removed a lot of passive voice, cleaned up wording, added some information.

Like any other microscopic traffic simulator, AORTA needs maps to run
simulations. One of AORTA's key features is that it generates maps using road data available from OpenStreetMap
(OSM) \cite{osm}, a Wikipedia-like interface for world maps. Researchers using
AORTA can export the relevant map area from OSM that is then parsed by AORTA to
create scale simulations of the real world. Using real world road data is
advantageous for a number of reasons. First, researchers using AORTA for human
traffic simulations would find it advantageous to also have the ability to use
real-world roads. Second, even as infrastructural changes will be made to the
road infrastructure to accommodate driver-less cars, it is reasonable to assume
that the \textit{same} road network will still be in use.\commentp{I
  don't understand this second point} Third, using
real-world data from OSM allows setting up simulations for an area of interest
in a matter of minutes, thereby allowing easy transfer of research results to
any desired region.

\commentd{The first reason for real world data almost feels too obvious to state
like that}
\commentp{I think all 3 reasons are similar.  I'd just say simplty
  that AORTA's key features are being open source, easy to install and
  get running, and flexibility to quickly and easily incorporate
  models from real road networks around the world.}

@@

AORTA is divided into three modular components: the map model, micro-simulation
engine, and UI. The map model transforms OSM maps into AORTA graphs, then
answers pathfinding and geometry queries. The simulation engine adds a notion of
agents, vehicle dynamics, and collisions. Finally, the UI, which uses Swing, can
interactively render the map and agents. A headless mode also enables running
experiments without the overhead of visualization.

@@

Our goal is to simulate traffic on existing infrastructure, not contrived,
randomly generated maps. Utilizing OpenStreetMaps, a community project offering
user-editable, free XML-based maps, enables this. To run a simulation on a map
from OSM, there is a one-time conversion process to transform the map into
AORTA's format, which is also encoded with XML.

@@

A map is represented as a directed graph with \emph{lanes} as edges and
\emph{turns} as vertices. A \emph{road} groups one or two groups of lanes
together, each group being a different direction of travel.
\emph{Intersections} map the possible turns from incoming to outgoing lanes.
Since both lanes and turns support traffic, they are grouped together as
\emph{traversables}, meaning agents may exist on them.

@@

Map construction is split into several passes, each feeding the next.
\begin{enumerate}
  \item OSM encodes many paths besides driveable roads, so these are first
        filtered out. GPS locations common to multiple roads are marked, since
        they implicitly indicate intersections.
  \item OSM \emph{ways} (sequences of GPS locations) include many
        intersections, so the builder next splits ways into undirected
        segments of roads between exactly two intersections.
  \item Undirected roads are multiplied into directed lanes in each direction
        (unless the road is marked one-way). The number of lanes is guessed from
        OSM's ``road type'' tag or an explicit number of lanes, when given.
  \item Turns are next constructed between incoming and outgoing lanes at each
        vertex. Left, right, straight, and U-turns are heuristically chosen based on
        the angles between lanes. When several lanes all cross into fewer lanes,
        merging is arbitrarily forced on the rightmost lanes.  Although inaccurate,
        these patterns are the best compromise. At the time of writing, even sources
        like Google Maps \cite{googleMaps} lack this information.
  \item Because turns are heuristically placed, the graph is usually
        disconnected. Since the largest resulting map is desirable, the strongly
        connected components of the digraph with the fewest members are removed.
        Disconnected lanes are spliced out of larger roads, leaving unappealing
        geometric gaps, but correctly ensuring connectivity for corner-case free
        pathfinding.
  \item The builder serializes the graph in a simple XML format.
\end{enumerate}

@@

\begin{enumerate}
  \item Introducing new agents into the map
  \item Updating the position and velocity of agents based on their choice the
        previous step
  \item Checking for collisions
  \item Allowing each agent to react
\end{enumerate}

Agents have a few enforced constraints -- they cannot exceed their physical
acceleration capabilities, travel in reverse, or travel past the end of a
traversable without appearing on the next. Intersection policies and avoiding
collisions on the same traversable provide the other constraints, but the
agent's behavior must satisfy these. There are three main components of the
simulation, each configurable by extending built-in implementations of an
interface: agent behaviors, routing strategies, and intersection policies.

Currently lane-changing is not supported. Merging lanes is accomplished by
treating the merge point as an intersection.

@@

To introduce new agents into the simulation in real-time, users can create
\emph{generators} programatically or using the UI. A generator picks a random
starting position and goal from two sets of roads. These roads can be selected
by mouse with a bounding polygon in the UI, letting specific traffic patterns
like rush-hour scenarios easily be constructed between two regions. A generator
may also cover the entire map, uniformly distributing traffic everywhere.

@@

Once an agent's route is ready, the agent waits alongside its starting lane (as
if it is in a driveway or parking lot). When the lane has no agents that could
potentially crash with the new vehicle, the new agent enters the system with
zero initial speed.  Because behaviors must enforce speed limit along a lane 
from the moment an agent enters the road, then if there is no agent within the
worst-case distance (traveling distance at the speed limit, plus stopping
distance from this speed) before the spawn point, then the new agent may enter.
To avoid predicting which agents might enter the lane in the near future, there
is a buffer space at the beginning of the lane where a new agent may never
spawn. This disqualifies some short lanes from being spawn points, but this is
not a big limitation in practice.

@@

All traversables (lanes and the straight line-approximation of turns) maintain
an ordered queue of occupying agents. One simulation step happens atomically,
eaning collision checking occurs after all agents have moved. If this was not
done, false positives could occur when an agent steps before the agent in front
of it. A collision is defined as the reversal of order of agents before and
after a step.

Likewise, intersections check for collisions by examining the turn each agent is
performing and verifying that no other agent is simultaneously performing a
conflicting turn. AORTA has a low-granularity model of turn conflict. If two
vehicles at any point along two different turns could ever come into contact,
then the turns are always in conflict, no matter where the agent specifically is
along that turn. The alternative of tiling intersections \cite{JAIR08-dresner}
would provide more granularity, but it adds computational complexity and would
be impractical for many of the complex intersection geometries inferred from
OSM.

@@

Each agent has a \emph{behavior} governing it. When the agent travels past the end
of a lane during its step, the behavior picks the turn to pursue. Each step,
the behavior can make the agent react by performing one of two actions:
disappearing from the map (only when the agent is at rest and is done with its
route) or setting an acceleration for the next step. Were lane changing or
merging modeled, that would be another possible action.

Currently, the primary behavior used by all agents is a generalized, baseline
behavior that guarantees not to collide with another agent or enter an
intersection at the wrong time. Since it picks the highest safe choice of
acceleration at each step, it could be easily extended to mimic human drivers by
traveling at some random slower acceleration or to optimize fuel efficiency
by tuning movement.

@@

The default behavior maximizes acceleration while obeying speed limits,
avoiding collisions, and obeying intersections. By considering the highest speed
achievable at the end of the next step and the subsequent stopping distance at
this speed, the behavior can scan sufficiently far ahead along several
traversables to locate agents to avoid and intersections to query. The analysis
short-circuits at the first nearby agent and intersection requiring a stop,
since it would be illegal to pass another anyway.

@@

Picking accelerations that optimize objective functions is a separate task from
picking a path to travel. Agent behaviors consult a route strategy to pick
turns. Currently two simple implementations are available. One statically plans
a route to some goal using standard A* search \cite{astar}; the other performs a
drunken walk, lazily planning random choices at each intersection as the
lookahead engine needs to know. Future experiments in dynamic replanning or
hierarchial planning can be performed without modifying any code other than a
creating new route implementation. Route policies could also interact with a
singleton overseer to gain some sort of global insight or with smart
intersections to avoid congestion.

@@

However, with very high volumes of traffic, it is possible for waiting agents to
accumulate and fill lanes to their capacity. When this happens in a circular
manner so that the agent at the front of each lane is waiting to perform a turn
into another full lane with the same such head agent, gridlock \cite{gridlock}
occurs; no agent in the system will make progress unless an agent at the front
decides to pick a different turn. At best, AORTA can detect this by maintaining
a dependency graph that encodes the ``is following'' relation and finding
cycles. It is not clear how to prevent or mend the deadlock without forcing
agents to re-route.

@@

AORTA is designed for controlled experiments. Consequently, simulations are
deterministically reproducible. A seed for the pseudo-random number generator,
the input graph, and the generators' configuration determine the outcome of any
simulation. This determinism may break down when generators use worker threads,
since routes may be computed in different orders, causing agents to enter the
road at different times.

@@

\commentd{http://sumo.sourceforge.net/pdf/mieec1.pdf gives their measure as
  agent updates / second. This is indep of dt and num of agents, assuming
  nothing grows nonlinearly with more agents. our initial result on my box is
  about 150,000 moves / second. so with 500 agents, runs 30x faster than
  reality. with 5,000, runs about 3x faster. with 50,000 (yes, this works!),
  runs at 1/3 the speed of reality (it takes 1 real second to simulate 0.3
  seconds)}

@@

This paper has presented AORTA, a new city-scale traffic simulation framework
that focuses on ease-of-use and policy configurability. In conjunction with Open
Street Maps, AORTA allows researchers to repeat an experiment in any number of
places with minimal effort. Future work will exploit the flexible infrastructure
of interactions between agent behaviors and intersections by exploring dynamic
replanning and adaptable signal timings. On top of these structures,
applications could experiment with dynamic contra-flow \cite{ITSC11-hausknecht}
and micro-tolling \commentd{need a cite}. There will also be an effort to
improve AORTA as a framework by including a lane-changing model and preventing
gridlock.

@@

\commentd{Er, so how do I finally end this?}

>>>>

Like any other microscopic traffic simulator, AORTA needs maps to run
simulations. One of AORTA's key features is that it generates maps using road data available from OpenStreetMap
(OSM) \cite{osm}, a Wikipedia-like interface for world maps. Researchers using
AORTA can export any relevant map area from OSM. This map is then parsed by AORTA to
create scale simulations of the real world. A researcher employing AORTA can have a simulation
of any part of the world for which OSM has data running in about five minutes
(allowing for download time).  AORTA also has the advantage of being open source and
easy to install so that researchers can make any tweaks they find necessary.
\footnote{http://code.google.com/p/road-rage}

@@

AORTA is divided into three modular components: the map model, micro-simulation
engine, and user interface (UI). The map model transforms OSM maps into AORTA graphs, then
answers pathfinding and geometry queries. The simulation engine adds a notion of
agents, vehicle dynamics, and collisions. Finally, the UI can
interactively render the map and agents. A headless mode also enables running
experiments without the overhead of visualization.

@@

One advantage of AORTA is that it simulates traffic on existing infrastructure rather than contrived,
randomly generated maps. Utilizing OpenStreetMaps, a community project offering
user-editable, free XML-based maps, enables this. To run a simulation on a map
from OSM, there is a one-time conversion process to transform the map into
AORTA's format, which is also encoded with XML.

@@

A map is represented as a directed graph with \emph{lanes} as edges and
\emph{turns} as vertices. A \emph{road} groups one or two sets of lanes
together, one set for each direction of travel.
\emph{Intersections} map the possible turns from incoming to outgoing lanes.
Since both lanes and turns support traffic, they are grouped together as
\emph{traversables}, meaning agents may exist on them.

@@

Map construction is split into several passes, each feeding the next.
\begin{enumerate}
  \item OSM encodes many paths besides driveable roads, so the builder first filters
        these out. Next the builder marks GPS locations common to multiple roads, since
        they implicitly indicate intersections.
  \item OSM \emph{ways} (sequences of GPS locations) include many
        intersections, so the builder next splits ways into undirected
        segments of roads between exactly two intersections.
  \item The builder multiplies undirected roads into directed lanes in each direction
        (unless the road is marked one-way). The builder guesses the number of lanes based on
        OSM's ``road type'' tag or an explicit number of lanes, when that data is available.
  \item The builder constructs turns between incoming and outgoing lanes at each
        vertex. The builder uses an angle heuristic to choose between left, right, straight,
        and U-turns. When several lanes all cross into fewer lanes,
        the builder forces merging of the rightmost lanes.  Although inaccurate,
        these patterns are the best compromise. At the time of writing, even sources
        like Google Maps \cite{googleMaps} lack this information.
  \item Because turns are heuristically placed, the graph is often
        disconnected. In order to ensure that an agent will be able to reach any point in the map,
        the builder removes all but the largest strongly connected component of the digraph.
        Disconnected lanes are spliced out of larger roads, leaving unappealing
        geometric gaps, but correctly ensuring connectivity for corner-case free
        pathfinding.
  \item The builder serializes the graph in a simple XML format.
\end{enumerate}

@@

\begin{enumerate}
  \item Introduce new agents into the map
  \item Update the position and velocity of agents based on their choice of
  acceleration in the previous step
  \item Check for collisions
  \item Allow each agent to choose a reaction for the next step
\end{enumerate}

Agents have a few enforced constraints -- they cannot exceed their physical
acceleration capabilities, travel in reverse, or travel past the end of a
traversable without appearing on the next. Intersection policies and avoiding
collisions on the same traversable provide the other constraints, but the
agent's behavior is responsible for obeying these. There are three main components of the
simulation, each configurable by extending built-in implementations of an
interface: agent behaviors, routing strategies, and intersection policies.

Currently agents cannot switch lanes. AORTA allows merging lanes by
treating the merge point as an intersection.

@@

To introduce new agents into the simulation in real-time, users can create
\emph{generators} programatically or using the UI. A generator picks a random
starting position and goal from two sets of roads. Users can select these roads 
by mouse with a bounding polygon in the UI, allowing specific traffic patterns
like rush-hour scenarios to be easily simulated. A generator
may also cover the entire map, uniformly distributing traffic everywhere.

@@

Once an agent's route is ready, the agent waits alongside its starting lane (as
if it is in a driveway or parking lot). When the lane has no agents that could
potentially crash with the new vehicle, the new agent enters the system with
zero initial speed.  Because behaviors must enforce speed limit along a lane 
from the moment an agent enters a road, it is safe for a new agent to enter a road
if there is no other agent within the worst-case distance (traveling distance at the
speed limit, plus stopping distance from this speed) preceeding the spawn point.
To avoid predicting which agents might enter the lane in the near future, there
is a buffer space at the beginning of the lane where a new agent may never
spawn. This disqualifies some short lanes from being spawn points, but this is
not a big limitation in practice.

@@

All traversables (lanes and the straight line-approximation of turns) maintain
an ordered queue of occupying agents. One simulation step happens atomically,
meaning collision checking occurs after all agents have moved. If this were not
the case, false positives could occur when an agent steps before the agent in front
of it. A collision is defined as the reversal of order of agents during a step.

Likewise, intersections check for collisions by examining the turn each agent is
performing and verifying that no other agent is simultaneously performing a
conflicting turn. AORTA has a low-granularity model of turn conflict. If two
vehicles at any point along two different turns could ever come into contact,
then the turns are always in conflict, no matter where the agents actually are
along that turn. The alternative of tiling intersections \cite{JAIR08-dresner}
would provide more granularity, but it adds computational complexity and would
be impractical for many of the complex intersection geometries inferred from
OSM.

@@

Each agent has a \emph{behavior} governing it. When the agent travels past the end
of a lane during its step, the behavior picks the turn to take. At each step,
the behavior can make the agent react by performing one of two actions:
disappearing from the map (when the agent is at rest and is done with its
route) or setting an acceleration for the next step. Were lane changing or
merging modeled, that would be another possible action.
\commentm{What's the difference between lane changing and merging?}

Currently, all agents use a primary behavior that is a generalized, baseline
behavior that guarantees not to collide with another agent or enter an
intersection at the wrong time. Since it picks the highest safe choice of
acceleration at each step, it could easily be extended to mimic human drivers by
traveling at some random lesser acceleration or to optimize fuel efficiency
by tuning movement.

@@

The default behavior maximizes acceleration while obeying speed limits,
avoiding collisions, and obeying intersections. By considering the highest speed
achievable at the end of the next step and the subsequent stopping distance at
this speed, the behavior can scan sufficiently far ahead along several
traversables to locate agents to avoid and intersections to query. The analysis
short-circuits at the first nearby agent and intersection requiring a stop,
since it would be illegal to drive further anyway.

@@

Picking accelerations that optimize objective functions is a separate task from
picking a path to travel. Agent behaviors consult a route strategy to pick
turns. Currently two simple implementations are available. One statically plans
a route to some goal using standard A* search \cite{astar}. The other performs a
drunken walk, lazily planning random choices at each intersection as the
lookahead engine requires, choosing the direction that minimizes the Euclidean
distance to the goal with 75\% probability. Researchers can experiment with
dynamic replanning or hierarchial planning without modifying any code other than
creating a new route implementation. Route policies could also interact with a
singleton overseer to gain some sort of global insight or with smart
intersections to avoid congestion.

@@

However, with very high volumes of traffic, it is possible for waiting agents to
accumulate and fill lanes to their capacity. When this happens in a circular
manner so that the agent at the front of each lane is waiting to perform a turn
into another full lane with the same such head agent, gridlock \cite{gridlock}
occurs; no agent in the system will make progress unless an agent at the front
decides to pick a different turn. At best, AORTA can detect this by maintaining
a dependency graph that encodes the ``is following'' relation and finding
cycles. It is not clear how to prevent or mend deadlock without forcing
agents to re-route.

@@

AORTA is designed for controlled experiments. Consequently, simulations are
deterministically reproducible. A seed for the pseudo-random number generator,
the input graph, and the generators' configuration fully determine the outcome of any
simulation. This determinism may break down when generators use worker threads,
since routes may be computed in different orders, causing agents to enter the
road at different times.

@@

\commentd{http://sumo.sourceforge.net/pdf/mieec1.pdf gives their measure as
  agent updates / second. This is indep of dt and num of agents, assuming
  nothing grows nonlinearly with more agents. our initial result on my box is
  about 150,000 moves / second. so with 500 agents, runs 30x faster than
  reality. with 5,000, runs about 3x faster. with 50,000 (yes, this works!),
  runs at 1/3 the speed of reality (it takes 1 real second to simulate 0.3
  seconds)}
\commentm{I typically have 160k-170k agent steps per second on my machine.  The
	highest I have seen is upwards of 300k, but that was because most agents had
	gridlocked...}

@@

This paper has presented AORTA, a new city-scale traffic simulation framework
that focuses on ease-of-use and policy configurability. In conjunction with Open
Street Maps, AORTA allows researchers to repeat an experiment in any number of
places with minimal effort. Future work will exploit the flexible infrastructure
of interactions between agent behaviors and intersections by exploring dynamic
replanning and adaptable signal timings. On top of these structures,
applications could experiment with dynamic contra-flow \cite{ITSC11-hausknecht}
and micro-tolling \commentd{need a cite}. There will also be an effort to
improve AORTA as a framework by including a lane-changing model and preventing
gridlock.
\commentm{Perhaps mention concurency plans (with TACC) to simulate arbitrarily large maps?}

@@

\commentd{Er, so how do I finally end this?}
\commentm{We still have a bit of passive voice.  How much is acceptable?  I think
	it sounds fine, but I imagine an English professor may disagree.}
